Data Specialization for a Staged Lambda Calculus

### 2015-02-26 paper/semantics.tex
edits to masking and top of dynamics

We evaluate \lang\ terms in two stages. The first stage of evaluation has 
three interdependent modes of operation, depending on the world of the term it is evaluating.
Specifically, first stage evaluation:
\begin{enumerate}
\item fully reduces terms in \bbonep, as they are purely first-stage;
\item reduces terms in \bbonem, save for any world \bbtwo\ terms they may
contain (inside \texttt{next}s);
\item scans terms in \bbtwo\ for any world \bbonem\ subterms (inside \texttt{prev}s), 
and reduces those in place.
\end{enumerate}
These tasks are accomplished by the $\redsym$, $\redonesym$, and $\redtwosym$
big-step judgments, respectively (defined in \ref{fig:diaSemantics}).

@@

We achieve this by threading through a \emph{residual table} $\xi$, which
functions as a top-level sequence of \texttt{let}-bindings of residuals $e$ to
variables $\hat{y}$ in an evaluated \bbonem\ term. Thus, $\redonesym$ evaluation
yields a pair $[\xi;v]$ of a residual table and a \emph{partial value} $v$
containing variables bound in that table. The grammar of residual tables and
partial values (terms fully evaluated at the first but not second stage) is
given in \ref{fig:values}.

@@

For all constructs other than $\prev{e}$, $\redtwosym$ evaluation proceeds
recursively on all subterms, yielding a monostaged residual $q$. At $\prev{e}$,
we evaluate $e$ using $\redonesym$ to $[\xi;v]$. Because $\colmix e{\fut A}$,
$v$ must have the canonical form $\next{\hat{y}}$ for some $\hat{y}$ bound in $\xi$. We \emph{reify}
$\xi$ into a sequence of second-stage \texttt{let} bindings around $\hat{y}$,
using the auxiliary judgment $\reify\xi q{q'}$ defined in
\ref{fig:diaSemantics}, then return the resulting residual $q'$.

>>>>

We evaluate \lang\ terms in two stages. The first stage of evaluation has 
three mutually recursive modes of operation, depending on the world of the term it is evaluating.
Specifically, first stage evaluation:
\begin{itemize}
\item fully reduces terms in \bbonep, as they are purely first-stage;
\item reduces terms in \bbonem, save for any world \bbtwo\ terms they may
contain (inside \texttt{next}s);
\item scans terms in \bbtwo\ for any world \bbonem\ subterms (inside \texttt{prev}s), 
and reduces those in place.
\end{itemize}
These tasks are accomplished by the $\redsym$, $\redonesym$, and $\redtwosym$
big-step judgments, respectively (defined in \ref{fig:diaSemantics}).

@@

We achieve this by threading through a \emph{residual table} $\xi$, which
functions as a top-level sequence of \texttt{let}-bindings of residuals $e$ to
variables $\hat{y}$ in an evaluated \bbonem\ term. Thus, $\redonesym$ evaluation
yields a pair $\rtab \xi v$ of a residual table and a \emph{partial value} $v$
containing variables bound in that table. The grammar of residual tables and
partial values (terms fully evaluated at the first but not second stage) is
given in \ref{fig:values}.

@@

For all constructs other than $\prev{e}$, $\redtwosym$ evaluation proceeds
recursively on all subterms, yielding a monostaged residual $q$. At $\prev{e}$,
we evaluate $e$ using $\redonesym$ to $\rtab \xi v$. Because $\colmix e{\fut A}$,
$v$ must have the canonical form $\next{\hat{y}}$ for some $\hat{y}$ bound in $\xi$. We \emph{reify}
$\xi$ into a sequence of second-stage \texttt{let} bindings around $\hat{y}$,
using the auxiliary judgment $\reify\xi q{q'}$ defined in
\ref{fig:diaSemantics}, then return the resulting residual $q'$.

### 2015-02-26 paper/splitting.tex
changed masking figure to use inference rules
edits to masking and top of dynamics

The point of the masking operation is simply to convert a partial value into a masked value $\mval i q$
by assigning all of the first-stage content to $i$ and all of the second-stage content to $q$.
The rules of the masking relation are given in \ref{fig:valMask}.

Masking operates by first inducting on the entries of the residual table.  
Being purely second-stage content, these are reified into let statements at the top of the resumer.
Once the table is empty, masking inducts on value itself.

Masking assigns values in \bbonep\ to the immediate value
and likewise assigns references into the residual table to the resumer.
In both cases, the alternate component is assigned to $\tup{}$, to represent trivial information.

@@

\input{figures/splitting-term}
\input{figures/splitting-misc}

>>>>

The masking operation converts a partial value into a masked value $\mval i q$,
where $i$ is a monostage value encoding all of the first-stage information of the original partial value,
and $q$ is a monostage term encoding all of the second-stage information of the original partial value.
Remember that masking will not play a role in the final splitting algorithm, 
but instead exists to specify the contract to which splitting at \bbonem\ must adhere.
Accordingly, masking should be thought of as a definition, and as such it is in some parts arbitrary.

The rules of the masking relation are given in \ref{fig:valMask}.
Masking operates by first inducting on the entries of the residual table.  
Being purely second-stage content, these are reified into let statements at the top of the resumer.
Once the table is empty, masking inducts on value itself.

Masking assigns ground values to the immediate value
and likewise assigns references into the residual table to the resumer.
In both cases, the alternate component is assigned to $\tup{}$, to represent trivial information.
Note that the \texttt{grnd} and \texttt{next} annotations are erased.

@@

\input{figures/splitting-misc}
\input{figures/splitting-term}
